---
title: "HiLo Exploratory Analysis"
author: "Fiona Francis"
date: "2/8/2021"
output: github_document
---
This analysis is for Sharon Jeffrey and is part of an overall project looking at the effect of current speed on fish and invertebrate communities at different sites in the Straight of Georgia. Jillian Campbell's MSc analysis was also a component of this project. 

Sharon is interested in using a mixed-effects model to look at how different factors including site current speed, depth and substrate type affect invert biomass and invert % cover. 

The study uses 14 sites, 7 "high" current sites and 7 "low" current sites. At each site Sharon collected invert data in 9 quadrats that are seperated into 3 depths (10, 30 and 50 ft depth). There is no repeat sampling so there are a total of 126 data points (14sites*9quadrats). However, because there are multiple collections at each site, a mixed effects model is necessary to account for replciation at the site level. I don't think that we need to nest transect within site because I think that each transect was at 1 depth (need to confirm this with Sharon) and so depth is already being included as a fixed effect and accounting for this replication (We need to worry about transect if they were run perpendicular to shore (i.e. from 10 to 50 m)).

For substrate, substrate 1 is >60% of quadrat and the codes are as follows: Substrate Codes: 1=Bedrock Smooth, 2=Bedrock w crevices, 3=Boulders, 4=Cobble, 5=Gravel, 6=Pea Gravel, 7=Sand, 9=Mud, 0=Wood/Bark, 10=Crushed Shell, 11=Whole/Chunk Shell

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Data

```{r, include = FALSE}
library(tidyverse)
library(fishualize)
library(nlme)
library(lme4)
library(bbmle)

invert <- read_csv("HiLoData_forFiona.csv")
site_info <- read_csv("sitelevel_data.csv")

head(invert)
head(site_info)
glimpse(invert)

# rename column in site_info

site_info <- site_info %>% rename(site_length_m = 3)

# look at which sites are which 
invert %>% filter(HiLo == "Hi") %>% distinct(Site)

# remove columns we don't need right now 

site_info <- site_info %>% select(Site, site_length_m, SlopeAngle, AvgDailMaxCurr, InvertSpRichness, AvgRkCov)

# add site length, slope and current speed to invert dataframe

all.data <- right_join(invert, site_info, by = "Site")
glimpse(all.data)
head(all.data)
print(all.data, n = 60)
all.data
```

## Data visualization

Hi vs Lo 

```{r, echo=FALSE}
ggplot(data = invert) + geom_point(aes(HiLo, Invert_Biomass, colour = HiLo, alpha = 1.4), position = position_jitter(w = 0.05, h = 0), size = 4, show.legend = F) +
  scale_color_fish_d(option = "Sparisoma_viride", direction = 1)
```


Hi vs Lo (Log10 Biomass)

```{r, echo=FALSE}
ggplot(data = invert) + geom_point(aes(HiLo, log10(Invert_Biomass), colour = HiLo, alpha = 1.4), position = position_jitter(w = 0.05, h = 0), size = 4, show.legend = F) +
  scale_color_fish_d(option = "Sparisoma_viride", direction = 1)
```

Biomass at each quad depth overall

```{r, echo=FALSE}
ggplot(data = invert) + geom_point(aes(Quadrat, Invert_Biomass, colour = Quadrat), position = position_jitter(w = 0.05, h = 0), size = 3, show.legend = F) +
   scale_color_fish_d(option = "Sparisoma_viride", direction = 1)
```

Biomass at quad depth by site

```{r, echo=FALSE}
ggplot(data = invert) + geom_point(aes(Quadrat, Invert_Biomass, colour = Quadrat),position = position_jitter(w = 0.05, h = 0), size = 3, show.legend = F) +
   scale_color_fish_d(option = "Sparisoma_viride", direction = 1) +
  facet_wrap(~ Site, nrow = 3)
```

Biomass at quad depth by site coloured by Hi Lo

```{r, echo=FALSE}
ggplot(data = invert) + geom_point(aes(Quadrat, Invert_Biomass, colour = HiLo), size = 3, show.legend = F) +
   scale_color_fish_d(option = "Sparisoma_viride", direction = 1) +
  facet_wrap(~ Site, nrow = 3)

```

## Same plots but looking at % invert cover instead of biomass

```{r, echo=FALSE}

# Plot Hi vs Lo percent
ggplot(data = invert) + geom_point(aes(HiLo, Invert_percent, colour = HiLo, alpha = 2), position = position_jitter(w = 0.05, h = 0), size = 4, show.legend = F) +
  scale_color_fish_d(option = "Gramma_loreto", direction = 1)

#Plot percent at each quad depth overall

ggplot(data = invert) + geom_point(aes(Quadrat, Invert_percent, colour = Quadrat), position = position_jitter(w = 0.05, h = 0), size = 3, show.legend = F) +
   scale_color_fish_d(option = "Gramma_loreto", direction = 1)

# plot percent at quad depth by site

ggplot(data = invert) + geom_point(aes(Quadrat, Invert_percent, colour = Quadrat), size = 3, show.legend = F) +
   scale_color_fish_d(option = "Gramma_loreto", direction = 1) +
  facet_wrap(~ Site, nrow = 3)

# plot percent at quad depth by site coloured by Hi Lo

ggplot(data = invert) + geom_point(aes(Quadrat, Invert_percent, colour = HiLo), size = 3, show.legend = F) +
   scale_color_fish_d(option = "Gramma_loreto", direction = 1) +
  facet_wrap(~ Site, nrow = 3)

```

## Plots of biomass vs current speed

Biomass vs current speed coloured by hi lo

```{r, echo=FALSE}

ggplot(data = all.data) + geom_point(aes(AvgDailMaxCurr, Invert_Biomass, colour = HiLo, alpha = 2), size = 4, show.legend = T) +
  scale_color_fish_d(option = "Sparisoma_viride", direction = 1)
```

Log10Biomass vs current speed coloured by hi lo

```{r, echo=FALSE}

ggplot(data = all.data) + geom_point(aes(AvgDailMaxCurr, log10(Invert_Biomass), colour = HiLo, alpha = 2), size = 4, show.legend = T) +
  scale_color_fish_d(option = "Sparisoma_viride", direction = 1)
```

Biomass vs current speed coloured by quadrat

```{r, echo=FALSE}
ggplot(all.data) + geom_point(aes(AvgDailMaxCurr, log10(Invert_Biomass), colour = Quadrat, alpha = 2), size = 4) +
   scale_color_fish_d(option = "Sparisoma_viride", direction = 1)
```

Biomass vs rock cover

```{r, echo=FALSE}
ggplot(all.data) + geom_point(aes(AvgRkCov, log10(Invert_Biomass), colour = Quadrat), size = 4) +
    scale_color_fish_d(option = "Sparisoma_viride", direction = 1)

```

Biomass vs Slope

```{r, echo=FALSE}
ggplot(all.data) + geom_point(aes(SlopeAngle, Invert_Biomass, colour = Quadrat), size = 4) +
    scale_color_fish_d(option = "Sparisoma_viride", direction = 1)

```


Plot of fixed effects against each other

Avg rock and slope are at the site level
```{r, echo=FALSE}
ggplot(all.data) + geom_point(aes(SlopeAngle, AvgRkCov, colour = Quadrat), size = 4) +
    scale_color_fish_d(option = "Sparisoma_viride", direction = 1)

ggplot(all.data) + geom_point(aes(SlopeAngle, AvgDailMaxCurr, colour = Quadrat), size = 4) +
    scale_color_fish_d(option = "Sparisoma_viride", direction = 1)

ggplot(all.data) + geom_point(aes(AvgRkCov, AvgDailMaxCurr, colour = Quadrat), size = 4) +
    scale_color_fish_d(option = "Sparisoma_viride", direction = 1)

```
## mixed-effect models

```{r, include = TRUE}
head(all.data)
glimpse(all.data)

# using max current speed, fitting a random effect for site
# need to specify using ML instead of REML. Here is the logic from talking to Dave Iles a statistician at Environment Canada. "I think this is happening because restricted maximum likelihood (REML) is being used to fit the models, rather than ML.  Annoyingly, REML is the default method for fitting mixed-effect models in both the lme4 and nlme packages - and ML needs to be explicitly specified to fit and compare models with the same random effect structure but different fixed effects."

#single variable
Null <- lme(log10(Invert_Biomass) ~ 1, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm1 <-  lme(log10(Invert_Biomass) ~ HiLo, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm2  <- lme(log10(Invert_Biomass) ~ Quadrat, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm3 <- lme(log10(Invert_Biomass) ~ SlopeAngle, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm4 <-  lme(log10(Invert_Biomass) ~ AvgRkCov,
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm5 <-  lme(log10(Invert_Biomass) ~ AvgDailMaxCurr, 
            random = ~ 1 | Site, data = all.data, method = "ML")


#two variables
Lm6 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + Quadrat, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm7 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + SlopeAngle, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm8 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm9 <- lme(log10(Invert_Biomass) ~ Quadrat + SlopeAngle, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm10 <- lme(log10(Invert_Biomass) ~ Quadrat + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm11 <-  lme(log10(Invert_Biomass) ~ SlopeAngle + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")

#three variables

Lm12 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + Quadrat, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm13 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + Quadrat + SlopeAngle, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm14 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + Quadrat + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm15 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + SlopeAngle + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")

Lm16 <- lme(log10(Invert_Biomass) ~ Quadrat + SlopeAngle + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")

# four variables

Lm17 <- lme(log10(Invert_Biomass) ~ AvgDailMaxCurr + Quadrat + SlopeAngle + AvgRkCov, 
            random = ~ 1 | Site, data = all.data, method = "ML")


names <- c("Null", "Current", "Depth", "Slope", "Rock", "ContCurrent")

AICtab(Null, Lm1, Lm2, Lm3, Lm4, Lm5, Lm6, Lm7, Lm8,Lm9, Lm10, Lm11, Lm12, Lm13, Lm14, Lm15, Lm16, Lm17, base=TRUE, weights=TRUE, logLik=TRUE)

AICctab(Null, Lm1, Lm2, Lm3, Lm4, Lm5,Lm5b, base=TRUE, weights=TRUE, logLik=TRUE)

## let's look at Lm1 in more detail and look at the diagnostics
# model summary
Lm12

plot(Lm12)

#another way to plot this with ggplot
ggplot(data.frame(biomass=predict(Lm12,type="link"),pearson=residuals(Lm12,type="pearson")),
       aes(x=biomass,y=pearson)) +
  geom_point() +
  theme_bw()

#qqnorm plot not great, not terrible, could try and add a variance structure or log transform biomass

qqnorm(residuals(Lm12))

# variables to include as fixed effects are current speed, depth, Average rock cover
```

Testing with Lme4 because I am confused

``` {r lme4, include = F} 
# testing with lme4 because I am very confused 

Null <- lmer(log10(Invert_Biomass) ~ 1 + (1|Site), data = all.data)

lmer1 <- lmer(log10(Invert_Biomass) ~ HiLo + (1|Site), data = all.data)

lmer2 <- lmer(log10(Invert_Biomass) ~ AvgDailMaxCurr + (1|Site), data = all.data)

lmer3 <- lmer(log10(Invert_Biomass) ~ AvgRkCov + (1|Site), data = all.data)

lmer4 <-  lmer(log10(Invert_Biomass) ~ Quadrat + (1|Site), data = all.data)

lmer5 <-  lmer(log10(Invert_Biomass) ~ SlopeAngle + (1|Site), data = all.data)

AICtab(Null, lmer1, lmer2, lmer3, lmer4, lmer5, base=TRUE, weights=TRUE, logLik=TRUE)
